---
title: "Assignment 5 - Meta-analysis of pitch in schizophrenia"
author: "Riccardo Fusaroli"
date: "3/7/2017"
output: html_document
---

```{r setup}
library(pacman)
p_load(metafor)
library(lmerTest)

data = read.csv('Data.csv',header=T,sep=";")
```

# Building on the shoulders of giants: meta-analysis

## Questions to be answered

1. What is the current evidence for distinctive patterns of pitch mean and pitch sd in schizophrenia? Report how many papers report quantitative estimates, your method to analyze them, the estimated effect size of the difference (mean effect size and standard error for pitch mean, same for pitch sd) and forest plots representing it. 

```{r estimate effect sizes of mean}
#getting effect size of mean for each study
data_1=escalc('SMD', n1i=SampleSizeSchizo, n2i=SampleSizeContros, m1i=PitchMeanSchizo, m2i=PitchMeanControls, sd1i=PitchMeanSchizoSD, sd2i=PitchMeanControlsSD, data = data)

#renaem these 2 new columns
data_1 = plyr::rename(data_1, replace=c("yi"="yi_mean","vi"="vi_mean"))
```

```{r estimate effect sizes of sd}
data_1=escalc('SMD', n1i=SampleSizeSchizo, n2i=SampleSizeContros, m1i=PitchSDSchizo, m2i=PitchSDControls, sd1i=PitchSDSchizoSD, sd2i=PitchSDControlsSD, data = data_1)

#rename
data_1 = plyr::rename(data_1, replace=c("yi"="yi_sd","vi"="vi_sd"))
```

```{r estimate the overall effect}
#first for mean
m_mean <- lmer(yi_mean ~ 1 + (1 |Article) , weights = 1/vi_mean, data=data_1, control=lmerControl(check.nobs.vs.nlev="ignore", check.nobs.vs.nRE="ignore",calc.derivs = FALSE))
summary(m_mean)

#and SD
m_sd <- lmer(yi_sd ~ 1 + (1 |Article) , weights = 1/vi_sd, data=data_1, control=lmerControl(check.nobs.vs.nlev="ignore", check.nobs.vs.nRE="ignore"))
summary(m_sd)

#now redo this using rma (more fancy stats)
rma_mean = rma(yi=yi_mean, vi=vi_mean, data=data_1,slab=Article)
rma_mean

rma_sd = rma(yi=yi_sd,vi=vi_sd, data=data_1,slab=Article)
rma_sd
```

```{r forest plots 1}
forest(rma_mean,main="Effect of mean")


forest(rma_sd,main="Effect of standard deviation")
```


2. Do the results match your own analysis from Assignment 3? If you add your results to the meta-analysis, do the estimated effect sizes change? Report the new estimates and the new forest plots.

```{r get data out of schizo}
library(dplyr)

#import dataset from schizo folder
schizo=read.csv("schizo_data.csv")

#summarize by diagnosis - mean+sd and sd+sd
meta_data=schizo %>% group_by(Diagnosis) %>% summarise(mean_mean=mean(mean),mean_sd=sd(mean),sd_mean=mean(sd),sd_sd=sd(sd))

#get number of participants
s_sum = length(unique(schizo$ID[schizo$Diagnosis == "Schizophrenia"]))
con_sum = length(unique(schizo$ID[schizo$Diagnosis == "Control"]))

ass3 = data.frame(
  Article = "CogSci et.al.",
  Year= 2017,
  SampleSizeSchizo =s_sum,
  SampleSizeContros = con_sum,
  PitchMeanSchizo = meta_data$mean_mean[2],
  PitchMeanControls = meta_data$mean_mean[1],
  PitchMeanSchizoSD = meta_data$mean_sd[2],
  PitchMeanControlsSD = meta_data$mean_sd[1],
  PitchSDSchizo = meta_data$sd_mean[2],
  PitchSDControls = meta_data$sd_mean[1],
  PitchSDSchizoSD = meta_data$sd_sd[2],
  PitchSDControlsSD = meta_data$sd_sd[1]
  )

```


```{r add study to big df}
#get Cohen's d for our study
#mean
ass3=escalc('SMD', n1i=SampleSizeSchizo, n2i=SampleSizeContros, m1i=PitchMeanSchizo, m2i=PitchMeanControls, sd1i=PitchMeanSchizoSD, sd2i=PitchMeanControlsSD, data = ass3)

#renaem these 2 new columns
ass3 = plyr::rename(ass3, replace=c("yi"="yi_mean","vi"="vi_mean"))

#sd
ass3=escalc('SMD', n1i=SampleSizeSchizo, n2i=SampleSizeContros, m1i=PitchSDSchizo, m2i=PitchSDControls, sd1i=PitchSDSchizoSD, sd2i=PitchSDControlsSD, data = ass3)

#rename
ass3 = plyr::rename(ass3, replace=c("yi"="yi_sd","vi"="vi_sd"))

#put to the rest of the data
data_all = rbind(data_1,ass3)
```

```{r and make models again}
#first for mean
m_mean_2 <- lmer(yi_mean ~ 1 + (1 |Article) , weights = 1/vi_mean, data=data_all, control=lmerControl(check.nobs.vs.nlev="ignore", check.nobs.vs.nRE="ignore"))
summary(m_mean_2)

#and SD
m_sd_2 <- lmer(yi_sd ~ 1 + (1 |Article) , weights = 1/vi_sd, data=data_all, control=lmerControl(check.nobs.vs.nlev="ignore", check.nobs.vs.nRE="ignore"))
summary(m_sd_2)

#now redo this using rma
rma_mean_2 = rma(yi=yi_mean, vi=vi_mean, data=data_all,slab=Article)
rma_mean_2

rma_sd_2 = rma(yi=yi_sd,vi=vi_sd, data=data_all,slab=Article)
rma_sd_2

2#and now plots
forest(rma_mean_2,main="Effect of mean")
forest(rma_sd_2,main="Effect of standard deviation")

```


3. Assess the quality of the literature: report and comment on heterogeneity of the studies (tau, I2), on publication bias (funnel plot), and on influential studies.

```{r heterogenity}
rma_mean_2 #tau^2=0,021 (CIs:-0.034 0.0762) I^2=26,15% -report Q because of p-value
rma_sd_2 #tau^2=3.499 (CIs:1.4317 5.5673)  I^2=98,18%
```

```{r publication bias}
funnel(rma_mean_2, main = "Random-Effects Model",xlab = "Standardized Mean Difference")
regtest(rma_mean_2) #not significant
ranktest(rma_mean_2) #not significant =>no evidence for publication bias

funnel(rma_sd_2, main = "Random-Effects Model",xlab = "Standardized Mean Difference") #this looks super bad
regtest(rma_sd_2) #significant
ranktest(rma_sd_2)
```


```{r influence of studies}
#for mean
inf_mean = influence(rma_mean_2)
print(inf_mean)
plot(inf_mean) 

inf_sd = influence(rma_sd_2)
print(inf_sd)
plot(inf_sd) 

#new data for new meta-analysis
data_2 = data_all

#remove Martinez for mean but not sd
data_2[10,13:14]=NA

#remove Cohen 2014 for sd - can delete whole row (it has NAs for mean)
data_2 = data_2[-9,]
```




## Tips on the process to follow:

- Download the data on all published articles analyzing pitch in schizophrenia: https://www.dropbox.com/s/pmgw8wzc308so1p/Data.csv?dl=0
- Following the procedure in the slides calculate effect size and standard error of the effect size per each study. N.B. we focus on pitch mean and pitch standard deviation.
 . first try using lmer (to connect to what you know of mixed effects models)
 . then use rma() (to get some juicy additional statistics)

- Build a forest plot of the results (forest(model))
 
- Go back to Assignment 3, add your own study to the data table, and re-run meta-analysis. Do the results change?

- Now look at the output of rma() and check tau and I2
